Ecological management of stochastic systems with long transients

### 2020-01-04 manuscript/manuscript.Rmd
remote dependencies

---
title: "Ecological management of stochastic systems with long transients"
date: "`r Sys.Date()`"
author:
  - name: Carl Boettiger
    email: cboettig@berkeley.edu
    affiliation: ucb
#    footnote: Corresponding Author
address:
  - code: ucb
    address: "ESPM Department, University of California, 130 Mulford Hall Berkeley, CA 94720-3114, USA"
abstract: |
  Long transients may be common in ecological dynamics, but the implications such dynamics have for ecological management have not been fully explored.  Long transient periods can easily be mistaken for stable-state dynamics, but may require dramatically different management policies.  Here, I explore the optimal management of stochastic ecological systems that may contain either a tipping point or a ghost attractor: a recently recognized mechanism which can give rise to long transients. I consider three approaches of increasing sophistication: (1) dynamic management under a fixed model, (2) management that accounts for uncertainty over possible models, and (3) adaptive management that actively learns the correct model over the management process. This analysis confirms the prediction that long transients can create considerable uncertainty and give rise to very different optimal management policies, but also illustrates that dynamic management that can either plan for this uncertainty or actively learn to decrease the uncertainty can promote successful management of long transients.
journal: Theoretical ecology
layout: 3p
bibliography: refs.bib
output: rticles::elsevier_article
preamble: |
  \journal{Theoretical Ecology}
  \usepackage{lineno}
  \linenumbers

>>>>

---
title: "Ecological management of stochastic systems with long transients"
date: "`r Sys.Date()`"
author:
  - name: Carl Boettiger
    email: cboettig@berkeley.edu
    affiliation: ucb
#    footnote: Corresponding Author
address:
  - code: ucb
    address: "ESPM Department, University of California, 130 Mulford Hall Berkeley, CA 94720-3114, USA"
abstract: |
  Long transients may be common in ecological dynamics, but the implications such dynamics have for ecological management have not been fully explored.  Long transient periods can easily be mistaken for stable-state dynamics, but may require dramatically different management policies.  Here, I explore the optimal management of stochastic ecological systems that may contain either a tipping point or a ghost attractor: a recently recognized mechanism which can give rise to long transients. I consider three approaches of increasing sophistication: (1) dynamic management under a fixed model, (2) management that accounts for uncertainty over possible models, and (3) adaptive management that actively learns the correct model over the management process. This analysis confirms the prediction that long transients can create considerable uncertainty and give rise to very different optimal management policies, but also illustrates that dynamic management that can either plan for this uncertainty or actively learn to decrease the uncertainty can promote successful management of long transients.
journal: Theoretical ecology
layout: 3p
bibliography: refs.bib
output: rticles::elsevier_article
keywords: 
  - transients, 
  - optimal control, 
  - adaptive management, 
  - stochasticity, 
  - uncertainty, 
  - ecological management
preamble: |
  \journal{Theoretical Ecology}
  \usepackage{lineno}
  \linenumbers
